---
title: "HW 10: Maps"
author: "Madison Hongsermeier-Baxter"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(raster)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```

# Part 1: Outline maps

## Example: States and counties

### Access specific states and store the results in three data frames.
```{r}
# Outline of the 48 contiguous states
usa <- map_data(map = "state") # Tidyverse or ggplot2 must be loaded

# Outline of just Missouri
us_missouri <- map_data(map = "state",
                        region = "missouri")

# Outline of Alabama, Florida, Mississippi, and North Carolina
us_other <- subset(x = usa, 
                   subset = region %in% c("alabama",
                                          "florida",
                                          "mississippi",
                                          "north carolina"))
```

### First layer of the map. These states are filled with gray.
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray")
```

### Add Missouri. This state is filled with black.
```{r}
 ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black")
```

### Add the outline of the states.
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black")
```

### Replace default theme and add coord_fixed layer.
```{r}
# Add the filled states first so that the black outlines of all
# states are overlaid on top.
us_map <- ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black") + 
  theme_map() +
  coord_fixed(1.3)

us_map
```

### Create map of Missouri with the outlines of the counties and Scott County filled with black.
```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))

mo_map <- ggplot() + 
  geom_polygon(data = missouri, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_polygon(data = mo_scott, aes(x = long, y = lat, group = group), fill = "black") +
  theme_map() +
  coord_fixed(1.3)

mo_map
```

### Use patchwork to display maps side by side.
```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5,1))
```

## Outline maps: your turn

### US map
```{r}
us_illinois <- map_data(map = "state",
                        region = "illinois")

us_map <- ggplot() + 
  geom_polygon(data = us_illinois, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black") + 
  theme_map() +
  coord_fixed(1.3)

us_map
```

### Illinois map
```{r}
illinois <- map_data("county", "illinois")
il_madison <- subset(illinois, subregion %in% c("madison"))
il_iroquois <- subset(illinois, subregion %in% c("iroquois"))
il_hamilton <- subset(illinois, subregion %in% c("hamilton"))

illinois_map <- ggplot() + 
  geom_polygon(data = illinois, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_polygon(data = il_madison, aes(x = long, y = lat, group = group), fill = "black") +
  theme_map() +
  geom_polygon(data = il_iroquois, aes(x = long, y = lat, group = group), fill = "gray") +
  theme_map() +
  geom_polygon(data = il_hamilton, aes(x = long, y = lat, group = group), fill = "gray") +
  theme_map() +
  coord_fixed(1.3)

illinois_map
```

### Both maps

#### Birth state: Illinois, Birth county: Madison County, Middle initial county: Iriquois County, Last inital county: Hamilton County
```{r}
us_map + illinois_map + plot_layout(ncol = 2, widths = c(1.5,1))
```

# Part 2: Dot distribution maps

## Example: Distribution of Gigantopithecus and Ursus

### Set the seed.
```{r}
# Global vars -------------------------------------------------------------

# Restrict the map to the western US.
min_long <- -126
max_long <- -104
min_lat  <-   31
max_lat  <-   50

set.seed(9618973)
```

### Load data. Randomly sample data. Filter the data to restrict the distribution to the longitude and latitude variables. Select only the name, longitude and latitude columns.
```{r}
# Bigfoot data ------------------------------------------------------------

bigfoot <- read_csv("~/Documents/BI485/hongsermeier-baxter_madison/data/bigfoot.csv")

bigfoot <- filter(bigfoot,
                  long >= min_long & long <= max_long,
                  lat >= min_lat & lat <= max_lat) %>% 
  sample_n(300) %>% 
  mutate(name = "Gigantopithecus") %>% 
  select(name, long, lat)
```

### Read in the bears.csv data.
```{r}
bears <- read_csv("~/Documents/BI485/hongsermeier-baxter_madison/data/bears.csv")

both_species <- bind_rows(bigfoot, bears)

head(both_species)
```
```{r}
tail(both_species)
```

### Get terrain map.
```{r}
base = get_stamenmap(bbox = c(min_long, 
                              min_lat, 
                              max_long, 
                              max_lat),
                     zoom = 4, 
                     maptype = "terrain-background")



ggmap(base)
```

### Plot the first layer. This layer shows the Bigfoot data.
```{r}
base <- ggmap(base)

base + geom_point(data = bigfoot,
                  aes(x = long,
                      y = lat))
```

### Plot the bear data.
```{r}
base + geom_point(data = bears,
                  aes(x = long,
                      y = lat))
```

### Plot both species on the map. Use aes to match a shape and fill color to each species.
```{r}
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2) +
  scale_shape_manual(values = c(22:23)) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")
```

## Distribution map: your turn

### Set seed.
```{r}
min_long <- -126
max_long <- -104
min_lat  <-   31
max_lat  <-   50

set.seed(1872353)
```

### Load data. Randomly sample data. Filter the data to restrict the distribution to the longitude and latitude variables. Select only the name, longitude and latitude columns.
```{r}
# Bigfoot data ------------------------------------------------------------

bigfoot <- read_csv("~/Documents/BI485/hongsermeier-baxter_madison/data/bigfoot.csv")

bigfoot <- filter(bigfoot,
                  long >= min_long & long <= max_long,
                  lat >= min_lat & lat <= max_lat) %>% 
  sample_n(300) %>% 
  mutate(name = "Gigantopithecus") %>% 
  select(name, long, lat)
```

### Read in the bears.csv data.
```{r}
bears <- read_csv("~/Documents/BI485/hongsermeier-baxter_madison/data/bears.csv")

both_species <- bind_rows(bigfoot, bears)

head(both_species)
```
```{r}
tail(both_species)
```

### Get terrain map.
```{r}
base = get_stamenmap(bbox = c(min_long, 
                              min_lat, 
                              max_long, 
                              max_lat),
                     zoom = 4, 
                     maptype = "terrain-background")



ggmap(base)
```

### Plot the first layer. This layer shows the Bigfoot data.
```{r}
base <- ggmap(base)

base + geom_point(data = bigfoot,
                  aes(x = long,
                      y = lat))
```

### Plot the bear data.
```{r}
base + geom_point(data = bears,
                  aes(x = long,
                      y = lat))
```

### Plot both species on the map. Use aes to match a shape and fill color to each species.

#### The Bigfoot layer is slightly different from the original but the bear distribution is the same as the original.
```{r}
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2) +
  scale_shape_manual(values = c(22:23)) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")
```

# Part 3: Bathymetry

## Example: Fiji earthquakes

### Define the longitude and latitude range for the map. Convert quake depth to feet.
```{r}
min_long <- -170
max_long <- 164
min_lat  <- -42
max_lat  <- -8

# This converts depth in km to depth in feet.
# Necessary because bathymetry data will be in feet
earthquakes <- quakes %>% 
   mutate(depth = depth * 3280.84)
```

### Obtain the data directly from NOAA.
```{r}
# Will auto read the saved file if present.

fiji <- getNOAA.bathy(lon1 = min_long,
                      lon2 = max_long,
                      lat1 = min_lat,
                      lat2 = max_lat,
                      antimeridian = TRUE,
                      keep = TRUE)
```

### Use marmap package to automatically plot bathymetric data with ggplot2.
```{r}
autoplot.bathy(fiji, 
         geom = c("raster", "contour"),
         size = 0.1,
         na.rm = TRUE)
```

### Use a custom gradient by adding a scale_fill_gradient2 layer. Label the x- and y- axes.
```{r}
# Using autoplot, which "knows" to use autoplot.bathy
autoplot(fiji, 
         geom = c("raster","contour"),
         size = 0.1,
         na.rm = TRUE) +
  scale_fill_gradient2(low = "dodgerblue4", 
                       mid = "gainsboro", 
                       high = "darkgreen",
                       name = "Depth (ft)") +
  labs(x = "Longitude",
       y = "Latitude") +
  theme(axis.title.y = element_text(angle = 0, 
                                    vjust = 0.5))
```

### Save the autoplot result to an object called base_map, and then add the Fiji earthquakes data using familiar ggplot layers. Use geom_point to plot the location of each seismic event.
```{r}


base_map <- autoplot(fiji, 
                     geom = c("raster","contour"),
                     size = 0.1,
                     na.rm = TRUE) +
  scale_fill_gradient2(low = "dodgerblue4", 
                       mid = "gainsboro", 
                       high = "darkgreen",
                       name = "Depth (ft)") +
  labs(x = "Longitude",
       y = "Latitude") +
  theme(axis.title.y = element_text(angle = 0, 
                                    vjust = 0.5))
```
```{r}
base_map + 
  geom_point(data = earthquakes,
             aes(x = long,
                 y = lat))
```

### Adjust the size of the points to reflect the magnitude of the each quake. Use the alpha argument to make the points more transparent.
```{r}
base_map + 
  geom_point(data = earthquakes,
             aes(x = long,
                 y = lat,
                 size = mag),
             alpha = 0.4)
```

### Use the scale_size_continuous layer to set the minimum and maximum point sizes used for magnitude.
```{r}
base_map + 
  geom_point(data = earthquakes,
             aes(x = long,
                 y = lat,
                 size = mag),
             alpha = 0.4) +
  scale_size_continuous(range = c(0.5,3), name = "Magnitude")
```

## Bathymetry: now you try it

```{r}
min_long <- -90
max_long <- -58
min_lat  <- 8
max_lat  <- 28
```

```{r}
blennies <- read_csv("~/Documents/BI485/hongsermeier-baxter_madison/data/blennies.csv")
```

```{r}
carib_sea <- getNOAA.bathy(lon1 = min_long,
                      lon2 = max_long,
                      lat1 = min_lat,
                      lat2 = max_lat,
                      antimeridian = FALSE,
                      keep = TRUE)
```

```{r}
base_map <- autoplot(carib_sea, 
                     geom = c("raster","contour"),
                     size = 0.1,
                     na.rm = TRUE) +
  scale_fill_etopo(guide = FALSE) +
  labs(x = "Longitude",
       y = "Latitude") +
  theme(axis.title.y = element_text(angle = 0, 
                                    vjust = 0.5))

base_map
```

### Add blennies.
```{r}
base_map + geom_point(data = blennies,
                      aes(x = long,
                          y = lat,
                          shape = name),
                      fill = "white",
                      size = 3)+
  scale_shape_manual(values = c(21:23))+
  labs(shape = "Species")+
  theme(legend.position = "bottom")+
  theme(legend.text = element_text(face = "italic"))
```






