---
title: "HW 06 Part 2: Wrangling"
author: "Madison Hongsermeier-Baxter"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(knitr)
install.packages("smatr")
library(smatr)
```

# darters

```{r}
file_path <- "~/Documents/BI485/hongsermeier-baxter_madison/data"
dat <- read.delim(file.path(file_path, "darters.txt"))
```

```{r}
start_pos <- c(1, 4, 15, 16, 17, 18, 21, 25, 29, 33, 36, 39, 40)
stop_pos <- c(3, 14, 15, 16, 17, 20, 24, 28, 32, 35, 38, 40, 41)
names <- c("specimen id", "species", "riffle", "age", "sex", "length", "depth", "velocity", "temperature", "majsub", "minsub", "majtype", "mintype")
darters <- read_fwf(file.path(file_path, "darters.txt"), fwf_positions(start_pos, stop_pos, col_names = names), skip = 15)
```

```{r}
darters
```

Use filter() to remove tetrazonum and zonale.

```{r}
darters2 <- darters %>%
  filter(!species %in% c("zonale", "tetrazonum"))
darters2
```

Sum together the majsub and minsub into a new column called total_substrate. Does total_substrate sum to 1 for all observations?

Yes, total_substrate sums to 1 for all observations.
```{r}
mutate(darters2,
  total_substrate = majsub + minsub
)
```

Change the riffle values of 1 and 2 to Riffle 1 and Riffle 2.
```{r}
darters2$riffle[darters2$riffle==1] <- "Riffle_1"
darters2$riffle[darters2$riffle==2] <- "Riffle_2"
darters2
```

Change the length data from centimeters to millimeters. It is more common to use mm for small fishes.
```{r}
darters2 <- mutate(darters2,
       length = length * 10)
```

```{r}
darters2
```

Change the sex values of f and m to Female and Male.
```{r}
darters2$sex[darters2$sex=="f"] <- "Female"
darters2$sex[darters2$sex=="m"] <- "Male"
darters2
```

Use group_by() and summarize() to summarize the mean length, depth, and velocity for each species for each riffle. Make a table with the summary means.

```{r}
darters_means <- darters2 %>% group_by(species) %>%
  summarise_at(.vars = vars(length, depth, velocity), 
               .funs = c(mean = "mean")) 
darters_means
```

Plot 1
Make a plot that uses the stat_summary() technique but use x = species and y = length. Add a geom_point() layer to add the mean length of each species to the plot. Use a filled square shape. You will have to use the summarized data set to get the mean for this geom.

```{r}
ggplot(data = darters2) +
  stat_summary(
    mapping = aes(x = species, y = length),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  ) +
  geom_point(data=darters_means, mapping = aes(x = species, y = length_mean), shape = 15)
```

Turn the plot so that the species are listed on the left axis and length on bottom axis
```{r}
ggplot(data = darters2) +
  stat_summary(
    mapping = aes(x = species, y = length),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  ) +
  geom_point(data=darters_means, mapping = aes(x = species, y = length_mean), shape = 15) +
  coord_flip()
```

Add facet_wrap() to separate the two riffles.
```{r}
ggplot(data = darters2) +
  stat_summary(
    mapping = aes(x = species, y = length),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  ) +
  geom_point(data=darters_means, mapping = aes(x = species, y = length_mean), shape = 15) +
  coord_flip() +
  facet_wrap(~ riffle, nrow = 2)
```

Add labs() so that length includes the appropriate unit of measurement in parentheses and that the species axis is not labeled (x = NULL).
```{r}
ggplot(data = darters2) +
  stat_summary(
    mapping = aes(x = species, y = length),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  ) +
  geom_point(data=darters_means, mapping = aes(x = species, y = length_mean), shape = 15) +
  coord_flip() +
  facet_wrap(~ riffle, nrow = 2) +
  labs(y = "Length (mm)", x = "Species")
```

How does the relationship of the median, mean, and range change between the two riffles? In other words, how does the size distribution change between the two riffles?

"Spectabile" has a similar range in both riffles, but a lower median in riffle 2. "Caeruleum" has a smaller range and lower median in riffle 2. "Blennioides" has a much larger range in riffle 2 and a slightly lower median. The mean is similar between rifles for all species.



Plot 2

Make a new plot with a geom_point() layer of length for each species
```{r}
ggplot(data = darters2, mapping = aes(x = species, y = length)) + 
  geom_point()
```

Change the point color to a lighter shade in the range of gray50 to gray70
```{r}
ggplot(data = darters2, mapping = aes(x = species, y = length)) + 
  geom_point(color = "gray50")
```

Add a stat_summary() layer with these arguments:
        fun.y = median
        geom = "point"
        size = 2.2
```{r}
ggplot(data = darters2, mapping = aes(x = species, y = length)) + 
  geom_point(color = "gray50") +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median,
    geom = "point",
    size = 2.2
  ) 
```

 Add another stat_summary(). Change the y function to mean, size to 3, and add the color = "15"maroon" argument.
```{r}
ggplot(data = darters2, mapping = aes(x = species, y = length)) + 
  geom_point(color = "gray50") +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median,
    geom = "point",
    size = 2.2
  ) +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = mean,
    size = 3,
    geom = "point",
    color = "maroon"
  )
```
 
Facet, label, and flip the graph as you did above
```{r}
ggplot(data = darters2, mapping = aes(x = species, y = length)) + 
  geom_point(color = "gray50") +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median,
    geom = "point",
    size = 2.2
  ) +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = mean,
    size = 3,
    geom = "point",
    color = "maroon"
  ) +
  coord_flip() +
  facet_wrap(~ riffle, nrow = 2) +
  labs(y = "Length (mm)", x = "Species")
```

# feathers



Use the locale = locale(<argument>) argument in the read_csv() function to specify the decimal mark.
```{r}
names <- c("capture_date", "day", "site", "urban_score", "cortF", "cortB", "sex", "smi", "tarsus", "mass", "basal_cort", "stress_cort")
file_path <- "~/Documents/BI485/hongsermeier-baxter_madison/data"
feathers <- read_csv(file.path(file_path, "feathers.csv"), locale = locale(decimal_mark = ","), skip = 1, col_names = names)
```

```{r}
feathers
```

Calculate the mean tarsus length of all individuals in the data set, and store it in a variable called mean_tarsus. You should know how to calculate the mean of all values in a column.
```{r}
mean_tarsus <- mean(feathers[["tarsus"]])
mean_tarsus
```

 Use mutate() to log-transform (log()) the body mass (mass) and the tarsus length (tarsus). Save these in new columns as log_mass and log_tarsus, respectively.
```{r}
feathers <- mutate(feathers,
  log_mass = log(mass),
  log_tarsus = log(tarsus)
)
```
 
  Run the following function major_axis <- sma(log_mass ~ log_tarsus, data = <data>). Substitute the name of your imported data for <data>. The sma() function calculates the standardized major axis regression.
```{r}
major_axis <- sma(log_mass ~ log_tarsus, data = feathers)
```

Apply the summary() funciton to major_axis.
        The value you want for b is the slope of the estimate.
```{r}
summary(major_axis)
```

Apply the coef() function to major_axis. The output is a named numeric vector with two elements. The value you want is the second element. Store that value in the variable b.
```{r}
coef(major_axis)
```

```{r}
b <- 1.864889
```

    Use mutate to calculate SMI and store the values in a new column called new_smi. Compare your new column with the column you imported. Your values should be nearly identical, differing by no more than 0.1.
```{r}
feathers <- mutate(feathers,
  new_smi = mass * (mean_tarsus / tarsus) ^ b
)
feathers
```
    
Group by site
```{r}
feathers %>% 
  group_by(site)
```

calculate the mean and standard error of the mean (SEY) for basal_cort and stress_cort
```{r}
mean_basal_cort <- mean(feathers$basal_cort, na.rm = TRUE)
mean_stress_cort <- mean(feathers$stress_cort, na.rm = TRUE)
sd_basal_cort <- sd(feathers$basal_cort, na.rm = TRUE)
sd_stress_cort <- sd(feathers$stress_cort, na.rm = TRUE)
se_basal_cort <- sd_basal_cort/(sqrt(43))
se_stress_cort <- sd_stress_cort/(sqrt(43))
```

Plot 1
    Make a boxplot of new_smi for each site. Do you see any clear differences among the four sites?
    
There are clear differences among the sites. "Niort" has the lowest "new_smi" values. "Larochelle" has higher values. "Villefollet" has higher values. "Cebc" has the highest values.
```{r}
ggplot(data = feathers, mapping = aes(x = new_smi, y = site)) + 
  geom_boxplot()
```

Plot 2 A common graph is to plot the means with error bars based on the standard errors of the means.
```{r}
min_basal_cort <- mean_basal_cort - se_basal_cort
max_basal_cort <- mean_basal_cort + se_basal_cort
min_basal_cort 
max_basal_cort
```

```{r}
ggplot(data = feathers, mapping = aes(x = site, y = basal_cort)) + 
  geom_point() +
  geom_errorbar(ymin = min_basal_cort, ymax = max_basal_cort)
```

Repeat the above plot but with two differences. Use the mean_stress and its standard error, and use geom_linerange() instead of geom_errorbar
```{r}
min_stress_cort <- mean_stress_cort - se_stress_cort
max_stress_cort <- mean_stress_cort + se_stress_cort
min_stress_cort 
max_stress_cort
```

```{r}
ggplot(data = feathers, mapping = aes(x = site, y = stress_cort)) + 
  geom_point() +
  geom_linerange(ymin = min_stress_cort, ymax = max_stress_cort)
```
