---
title: "HW 06 Part 1: Complete the sections"
author: "Madison Hongsermeier-Baxter"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

Pay careful attention to the useful creation functions in section 5.5.1. 

Do not answer any of the questions.

```{r}
library(tidyverse)
library(nycflights13)
```

This code uses "mutate()" to create new columns "gain" and "speed" and add them to displayed data. 
```{r}
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)
```

This code refers to the newly created columns, "gain" and "hours" in order to use "mutate()" and create "gain_per_hour".
```{r}
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

This code uses "transmute()" in order to display only the newly created columns "gain", "hours", and "gain_per_hour".
```{r}
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

This code uses "%/%" and "%%" to calculate "hour" and "minute" based on the information "dep_time".
```{r}
transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```

This code uses "lag(x)" and "lead(x)" to refer to leading or lagging values of "(x <- 1:10)".
```{r}
(x <- 1:10)
lag(x)
lead(x)
```

This code uses "cumsum()" and "cummean()" to find the cumulative sum and cumulative mean of "x".
```{r}
x
cumsum(x)
cummean(x)
```

This code uses "min_rank()" to rank data from "y" in a ascending order and "min_rank(desc())" to rank in descending order
```{r}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
min_rank(desc(y))
```

This code uses "row_number()", "dense_rank()", "percent_rank()", and "cume_dist()" as various ways to rank data in ascending order.
```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```

### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.



### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.

