---
title: "HW 07: Data Visualization"
author: "Madison Hongsermeier-Baxter"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(GGally)
library(patchwork)
```

# Part 1: Recreate the Zuur figures

## Import and Wrangle

Import sparrow.txt.
```{r}
file_path <- "~/Documents/BI485/hongsermeier-baxter_madison/data"
sparrows_raw <- read_tsv(file.path(file_path, "sparrows.txt"))
```

View() the imported data to be sure the data look like they were properly imported.
```{r}
View(sparrows_raw)
```

Add two columns. Create new tibble called sparrows. Use this tibble for the remaining steps.
```{r}
sparrows <- sparrows_raw %>% 
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))
```

Run View(sparrows) to ensure that you have both columns.
```{r}
View(sparrows)
```

## Figure 2: outliers

Generate boxplot.
```{r}
plot_box <- sparrows %>% 
  ggplot(aes(x = box_group, 
             y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs(x = NULL,
       y = "Wing length (mm)")
```

Generate Cleveland dot plot.
```{r}
plot_cleveland <- sparrows %>% 
  ggplot(aes(x = wingcrd,
             y = row_order)) +
  geom_point() +
  labs(x = "Wing length (mm)",
       y = "Order of the data")
```

Put the two plots side-by-side.
```{r}
plot_box + plot_cleveland
```

Switch x axis and y axis.
```{r}
plot_box + plot_cleveland + coord_flip()
```

## Figure 3: outliers

```{r}
p_culmen <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_culmen
```

```{r}
p_nalospi <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_nalospi
```

```{r}
p_wt <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_wt
```

```{r}
p_wingcrd <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_wingcrd
```

```{r}
p_tarsus <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = tarsus)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_tarsus
```

```{r}
p_head <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_head
```

Produce six graphs and lay them out as shown in the figure.
```{r}
p_culmen + p_nalospi + p_wt + p_wingcrd + p_tarsus + p_head + plot_layout(ncol = 3)
```

## Figure 5: distribution

histogram
```{r}
h1 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2))
h1
```

Filter out months other than 6, 7, and 8. Add month names. Facet by month.
```{r}
h2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2)) +
  facet_wrap(~ Month, ncol = 1)
h2
```

Display the two plots as a singke figure.
```{r}
h1 + h2
```

## Figure 10: Relationships

Use the columns argument of ggscatmat to specify the first seven columns of data.
```{r}
sparrows %>% ggscatmat(columns = 1:7)
```

## Figure 11: interactions

Months are converted to factors.
```{r}
month_order = c("May", "June", "July", "August", "September")
sparrows %>% 
  filter(Month %in% 5:9,
         Sex != 0) %>% 
  mutate(Month = case_when(
    Month == 5 ~ "May",
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August",
    Month == 9 ~ "September"),
    Month = factor(Month, levels = month_order, ordered = TRUE),
    Sex = ifelse(Sex == 4, "Male", "Female")) %>% 
  ggplot(aes(x = wingcrd,
             y = wt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(Sex ~ Month)
```

# Part 2: Habitat parameters for Aegla crabs

Import the raw data from aegla_crabs.csv.
```{r}
library(tidyverse)
file_path <- "~/Documents/BI485/hongsermeier-baxter_madison/data"
crabs <- read_csv(file.path(file_path, "aegla_crabs.csv"))
crabs
```

Use select() to remove the Site, AT, WT, Shrimp, Aeglam, AeglaFe, AeglaFo columns.
```{r}
crabs2 <- crabs %>%
select(Width, Depth, Flow, pH, TDS, Cond, N, Ni, Nt, Phos, Mg, Ca, Crab)
crabs2
```

Use drop_na() to remove the one row with missing water quality data. Create a row_order dummy variable like we did for the sparrows.
```{r}
crabs3 <- 
  crabs2 %>%
  drop_na() %>%
         mutate(row_order = 1:nrow(.))
crabs3
```

Explore data to find variables with outliers.

width
```{r}
width1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = Width)) 
```

depth
```{r}
depth1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = Depth))
```

flow
```{r}
flow1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = Flow))
```

pH
```{r}
ph1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = pH))
```

TDS
```{r}
tds1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = TDS))
```

cond
```{r}
cond1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = Cond))
```

N
```{r}
n1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = N))
```

Ni
```{r}
ni1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = Ni))
```

phos
```{r}
phos1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = Phos))
```

Mg
```{r}
mg1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = Mg))
```

Ca
```{r}
ca1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = Ca))
```

Crab
```{r}
crab1 <- crabs3 %>% 
  ggplot(aes(x = row_order)) +
  geom_boxplot(aes(y = Crab))
```

```{r}
width1 + depth1 + flow1 + ph1 + tds1 + cond1 + n1 + ni1  + phos1 + mg1 + ca1 + crab1 + plot_layout(ncol = 6)
```

There are three apparent outliers among the variables but only one that I think is an actual outlier. Find them and justify your choices.

Width, depth, and pH all have outliers. These variables all show single outliers with significant distance from the majority of the data. Width seems to have the most apparent outlier; it it the most distant outlier.


Choose four variables to make Cleveland plots, and make a 2x2 grid with the four plots, using patchwork.

```{r}
width2 <- crabs3 %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Width)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
n2 <- crabs3 %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = N)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
flow2 <- crabs3 %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Flow)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
ph2 <- crabs3 %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = pH)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
width2 + n2 + flow2 + ph2 + plot_layout(ncol = 2)
```

Choose three different variables to make histograms. Play with the bins or binwidths argument that you think reveals the data well. Save each histogram to a unique variable.

```{r}
hist_crab <- crabs3 %>%
  ggplot() +
  geom_histogram(aes(x = Crab),
                 binwidth = 1,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(0, 24, by = 2))
hist_crab
```

```{r}
hist_width <- crabs3 %>%
  ggplot() +
  geom_histogram(aes(x = Width),
                 binwidth = .5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(0, 12, by = 2))
hist_width
```

```{r}
hist_depth <- crabs3 %>%
  ggplot() +
  geom_histogram(aes(x = Depth),
                 binwidth = .5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(0, 5, by = 1))
hist_depth
```

Use the same three variables to produce density plots. Save each plot to a unique variable.

```{r}
density_crab <- crabs3 %>% 
  ggplot() +
  geom_density(aes(x = Crab)) +
  scale_x_continuous(breaks = seq(0, 24, by = 2))
density_crab
```

```{r}
density_width <- crabs3 %>% 
  ggplot() +
  geom_density(aes(x = Width)) +
  scale_x_continuous(breaks = seq(0, 12, by = 2))
density_width
```

```{r}
density_depth <- crabs3 %>% 
  ggplot() +
  geom_density(aes(x = Depth)) +
  scale_x_continuous(breaks = seq(0, 5, by = 1))
density_depth
```

Use `patchwork to produce a 2 column matrix with the histograms in the left column and the corresponding density plots in the right column.

```{r}
histograms <- hist_crab + hist_width + hist_depth + plot_layout(nrow = 3)
densityplots <- density_crab + density_width + density_depth + plot_layout(nrow = 3)
histograms | densityplots
```

Use ggpairs to explore the relationships among all the variables.

```{r}
crabs3 %>% ggpairs(columns = 1:14)
```
