---
title: "HW05 Part 2: Import"
author: Madison Hongsermeier-Baxter
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Limpets

```{r}
library(tidyverse)
file_path <- "~/Documents/BI485/hongsermeier-baxter_madison/data"
dat <- read_csv(file.path(file_path, "limpets.csv"))
```

```{r}
dat 
```

Use the col_types argument. The three column types are numeric, character, and character.
```{r}
read_csv(file.path(file_path, "limpets.csv"), col_types = c("ncc"))
```

Make the Sites column an ordered factor with these levels:

    PBL, KNRM, VBG, WP, PF, DP, SIO, CTZ1, CTZ2, CNM
```{r}
sites_ordered <- limpets.csv$Site <- factor(limpets.csv$Site, levels = c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO", "CTZ1", "CTZ2", "CNM"), ordered = TRUE)
```

```{r}
sites_ordered
```

Make a boxplot of length (in millimeters) for each site, colored by protected status.
```{r}
ggplot(data = limpets.csv, mapping = aes(x = Site, y = Length, color = Protected)) + 
  geom_boxplot()
```

Change the axis labels so they begin with capital letters. The y-axis should include the unit of measurement in parentheses
```{r}
ggplot(data = limpets.csv, mapping = aes(x = Site, y = Length, color = Protected)) + 
  geom_boxplot() +
  labs(x = "Site", y = "Length (mm)")
```

The boxplot shows that protected sites tend to have larger limpets. The boxplot also shows that sites "WP" and "DP" have outliers. 

# Roseate Terns

```{r}
file_path <- "~/Documents/BI485/hongsermeier-baxter_madison/data"
dat <- read.delim(file.path(file_path, "roseate_terns.txt"))
```


```{r}
dat
```

Tidy data
```{r}
tidy_roseate <- dat %>% 
  gather("AR", "AL", "AC", "BR", "BL", "BC", "JR", "JL", "JC", key = "Site", value = "Individuals")
tidy_roseate
```

 use filter to remove sites with missing counts
```{r}
filter(tidy_roseate, is.na(Individuals) == FALSE)
```

Make a line plot of population size over time.
```{r}
ggplot(data = tidy_roseate, mapping = aes(x = year, y = Individuals)) + 
  geom_line() +
  facet_wrap(~ Site, nrow = 2)
```

Which population(s) obviously increased in size between 1992 and 2016? Which population(s) obviously decreased in size during that time? Some lines have breaks in them. That is, they are not continuous across all years. Why?

The population at site "AR" most obviously increased during the time period. There also seem to be more slight increases in population at the sites "AC", "AL", "BC", and "JC". There were decreases at sites "BR" and "JR". Some lines have breaks in them due to missing values in the data.

# darters 

```{r}
file_path <- "~/Documents/BI485/hongsermeier-baxter_madison/data"
dat <- read.delim(file.path(file_path, "darters.txt"))
```

```{r}
start_pos <- c(1, 4, 15, 16, 17, 18, 21, 25, 29, 33, 36, 39, 40)
stop_pos <- c(3, 14, 15, 16, 17, 20, 24, 28, 32, 35, 38, 40, 41)
names <- c("specimen id", "species", "riffle", "age", "sex", "length", "depth", "velocity", "temperature", "majsub", "minsub", "majtype", "mintype")
darters <- read_fwf(file.path(file_path, "darters.txt"), fwf_positions(start_pos, stop_pos, col_names = names), skip = 15)
```

```{r}
darters
```

 Make riffle an unordered factor with levels 1 and 2.
```{r}
darters_riffle <- darters$riffle <- factor(darters$riffle, levels = c("1", "2"), ordered = FALSE)
```

Make major_type an ordered factor with levels s, fg, sg, lg, c
```{r}
darters_majtype <- darters$majtype <- factor(darters$majtype, levels = c("s", "fg", "sg", "lg", "c"), ordered = TRUE)
```

    Filter to remove rows with “zonale” and “tetrazonum”.
```{r}
darters2 <- darters %>%
  filter(!species %in% c("zonale", "tetrazonum"))
darters2
```

Remove mintype and minsub columns.
```{r}
darters3 <- subset(darters2, select = -c(mintype,minsub) )
darters3
```

Rename majsub and majtype to major_substrate and major_type, respectively.
```{r}
colnames(darters3)[colnames(darters3)=="majsub"] <- "major_substrate"
colnames(darters3)[colnames(darters3)=="majtype"] <- "major_type"
darters3
```

Data were collected from two riffles, separated by several hundred meters. The plots below explore differences between riffles. Use facet_wrap() to make pairs of plots separated by riffle.
Plot 1 Plot length as a function of depth. Map species to color and shape. What differences do you see between the two riffles?

Riffle 2 has higher length values. Riffle 1 has higher depth values. Riffle 2 has more blennioides.
```{r}
ggplot(data = darters3, mapping = aes(x = length, y = depth, color = species, shape = species)) + 
  geom_point() +
  facet_wrap(~ riffle, nrow = 2)
  
```


Plot 2: Make a boxplot of length for each of the three species. Which riffle shows the greatest number of outliers?

Riffle 2 has more outliers.
```{r}
ggplot(data = darters3, mapping = aes(x = species, y = length)) + 
  geom_boxplot() +
  facet_wrap(~ riffle, nrow = 2)
```

Plot 3: Make a boxplot of length for major substrate types for each species for each riffle. This will actually be six plots in one! To do this, use facet_grid(species ~ riffle) in place of facet_wrap(). How does the plot change if you switch the order of the argument (riffle ~ species) in the facet_grid() function?

```{r}

```

# abalone 

```{r}
file_path <- "~/Documents/BI485/hongsermeier-baxter_madison/data"
abalone <- read_csv(file.path(file_path, "abalone.csv"))
```

```{r}
abalone
```

remove the first column
```{r}
abalone1 <- select(abalone, 2:10)
abalone1
```

make a boxplot of height differences among the three types
```{r}
ggplot(data = abalone1, mapping = aes(x = Type, y = Height)) + 
  geom_boxplot() 
```

Make a scatterplot to see if height appears to correlate with rings
```{r}
ggplot(data = abalone1, mapping = aes(x = Rings, y = Height)) + 
  geom_point() 
```

Based on the graph, are the extraordinarily large individuals really old individuals?
No, based on the graphs, the outliers for size are not some of the older individuals according to the rings, an indicator of age.

Filter the data to remove the two large individuals
```{r}
abalone2 <- filter(abalone1, Height < 0.3)
abalone2
```

Change Type to an ordered factor
```{r}
abalone_type <- abalone2$Type <- factor(abalone2$Type, levels = c("I", "F", "M"), ordered = TRUE)

```

redo the scatterplot that you just made with the newly wrangled data
```{r}
ggplot(data = abalone2, mapping = aes(x = Rings, y = Height)) + 
  geom_point() +
  facet_wrap(~ Type, nrow = 2)
```

What patterns emerge? Which type is the largest? Are all females and males larger than immatures?

In general, females and males are larger than immatures, but there is a lot of overlap in size. 

Print the records of the individuals with zero height
```{r}
filter(abalone2, Height == 0)
```

assign NA (missing data) to those two records
```{r}
abalone2[abalone2 == 0] <- NA
```

Repolot the data to ensure the two observations are not included in the graph.
```{r}
abalone3 <- filter(abalone2, is.na(Height) == FALSE)
```

Make two scatterplots of your choice, between any two pairs of continuous variables that make sense to show as scatterplots. Color, shape, or both should distinguish the three types.

```{r}
ggplot(data = abalone3, mapping = aes(x = Diameter, y = Height, color = Type)) + 
  geom_point()
```

```{r}
ggplot(data = abalone3, mapping = aes(x = LongestShell, y = ShellWeight, color = Type)) + 
  geom_point()
```

