---
title: "HW05 Part 1: Complete the sections"
author: Madison Hongsermeier-Baxter
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 10: Tibbles

Read [R4ds Chapter 10: Tibbles](https://r4ds.had.co.nz/tibbles.html), sections 1-3.

### 10.1: Introduction

Load the tidyverse package. 

```{r}
library("tidyverse")
```


### 10.2: Creating tibbles

Enter your code chunks for Section 10.2 here.

Describe what each chunk code does. 


This code creates a tibble from a dataset. 
```{r}
as_tibble(iris)
```


This code creates a tibble from individual vectors.
```{r}
tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)
```


This code creates a tibble with non-syntatic column names.
```{r}
tb <- tibble(
  `:)` = "smile", 
  ` ` = "space",
  `2000` = "number"
)
tb
```


This code creates a tibble, using "tribble()" in order to lay out the data in an easily readable form.
```{r}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
)
```

### 10.3: Tibbles vs data.frame

Enter your code chunks for Section 10.2 here.

Describe what each chunk code does. 


This code creates a tibble for a large data set. Because the data set is so large, the tibble shows only 10 rows at a time.
```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
```


This code creates a tibble for a large data set. "n = 10" is used to specify that only 10 rows are displayed at once, and "width = Inf" is used to specify that all of the columns are displayed at once.
```{r}
nycflights13::flights %>% 
  print(n = 10, width = Inf)
```


This code creates a tibble that displays the entire data set.
```{r}
nycflights13::flights %>% 
  View()
```


This code selects a single variable from the data. "df$x" and "df[["x"]]" extract by name and "df[[1]]" extracts by position. 
```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)
df$x
df[["x"]]
df[[1]]
```


This code selects a single variable. "." is used as a placeholder.
```{r}
df %>% .$x
df %>% .[["x"]]
```

### 10.4: Not required

#### Section 10.5 Questions

Answer the questions *completely.* Use code chunks, text, or both, as necessary.

**1:** How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame). Identify at least two ways to tell if an object is a tibble. *Hint:* What does `as_tibble()` do? What does `class()` do? What does `str()` do?


There are multiple ways to tell if an object is a tibble. "class(mt_cars)" and "str(mt_cars)" both show that mt_cars is a data set, not a tibble.
```{r}
mtcars
```

```{r}
as_tibble(mtcars)
```

```{r}
class(mtcars)
```

```{r}
str(mtcars)
```

**2:** Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?


Some of the default data frame behaviors shown below may cause frustration. For example, with [] operations, data frames sometimes return data frames and sometimes return vectors. However, tibble will always return a tibble in response to a [] operation.
```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```

```{r}
tb <- tibble(abc = 1, xyz = "a")
tb$x
tb[, "xyz"]
tb[, c("abc", "xyz")]
```

## Chapter 11: Importing data

Read [R4ds Chapter 11: Data Import](https://r4ds.had.co.nz/data-import.html), sections 1, 2, and 5.

### 11.1 Introduction

Nothing to do here unless you took a break and need to reload `tidyverse`.

```{r}
library("tidyverse")
```

### 11.2 Getting started.

Do *not* run the first code chunk of this section, which begins with `heights <- read_csv("data/heights.csv")`. You do not have that data file so the code will not run.

Enter and run the remaining chunks in this section.


This code creates a tibble with column specification.
```{r}
read_csv("a,b,c
1,2,3
4,5,6")
```


This code creates tibbles where several lines are skipped. "skip = 2" specifies that the first 2 lines are skipped, and "skip = #" specifies that lines starting with "#" are skipped.
```{r}
read_csv("The first line of metadata
  The second line of metadata
  x,y,z
  1,2,3", skip = 2)

read_csv("# A comment I want to skip
  x,y,z
  1,2,3", comment = "#")
```


This code creates a tibble without column names. "col_names = FALSE" specifies that the first row should not be treated as headings.
```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```


This code creates a tibble. The data does not include column names, so "col_names = c("x", "y", "z"))" is used to specify the column names.
```{r}
read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```


This code creates a tibble that includes a missing value. "na = ." is used to specify the value that is being used to represent the missing value.
```{r}
read_csv("a,b,c\n1,2,.", na = ".")
```

#### 11.2 Questions

**1:** What function would you use to read a file where fields were separated with "|"?

You would use "read_delim()" and "delim = |"

**2:** (This question is modified from the text.) Finish the two lines of `read_delim` code so that the first one would read a comma-separated file and the second would read a tab-separated file. You only need to worry about the delimiter. Do not worry about other arguments. Replace the dots in each line with the rest of your code. 

# Comma-separated
`file <- read_delim("file.csv", delim = ",")`
# Tab-separated
`file <- read_delim("file.csv", delim = \t)`

**3:** What are the two most important arguments to `read_fwf()`? Why?

The most important arguments are "col_positions" and "col_types". These arguments are important to properly order the table.

**4:** Skip this question


**5: ** Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

In the first file, the header specifies 2 columns instead of 3, causing the third column to be dropped when you run the code.
In the second file, the header specifies 3 columns, but there are only 2 values in the first row, and there are 4 values in the second row. When you run the code, the first row has a missing value, and one of the values in the second row is dropped.
In the third file, there is only one value, so when you run the code, there is a missing value in the row.
In the fourth file, both columns have strings. When you run the code, both columns are treated as character vectors. 
In the fifth file, values are separated by ";" but they should be separated by ",". When you run the code, there appears to be only one column.

```{r}
read_csv2("a;b\n1;3")

```{r}
read_csv("a,b\n1,2,3\n4,5,6")
read_csv("a,b,c\n1,2\n1,2,3,4")
read_csv("a,b\n\"1")
read_csv("a,b\n1,2\na,b")
read_csv("a;b\n1;3")
```

### 11.3 and 11.4: Not required

### 11.5: Writing to a file

Just read this section. You may find it helpful in the future to save a data file to your hard drive. It is basically the same format as reading a file, except that you must specify the data object to save, in addition to the path and file name.

### 11.6 Not required

## Chapter 18: Pipes

Read [R4ds Chapter 18: Pipes](https://r4ds.had.co.nz/pipes.html), sections 1-3.

Nothing to do otherwise for this chapter. Is this easy or what?

**Note:** Trying using pipes for all of the remaining examples. That will help you understand them.

## Chapter 12: Tidy Data

Read [R4ds Chapter 12: Tidy Data](https://r4ds.had.co.nz/tidy-data.html), sections 1-3, 7. 

### 12.1 Introduction

Nothing to do here unless you took a break and need to reload the `tidyverse.`

### 12.2 Tidy data

Study Figure 12.1 and relate the diagram to the three rules listed just above them. Relate that back to the example I gave you in the notes. Bear this in mind as you make data tidy in the second part of this assignment.

You do not have to run any of the examples in this section.

### 12.3

Read and run the examples through section 12.3.1 (gathering), including the example with `left_join()`. We'll cover joins later.


This code represents a data set where some of the column names are values of a variable.
```{r}
table4a
```


This code is a version of the previous code that has been modified to tidy the data set by gathering the columns.
```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
```


This code uses "gather()" to tidy the data set "table4b".
```{r}
table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
```


This code combines the tidied versions of the data sets into one tibble.
```{r}
tidy4a <- table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
tidy4b <- table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
left_join(tidy4a, tidy4b)
```


This code represents a tibble in which the observations are spread across two rows.
```{r}
table2
```


This code is a modified version of the previous code that uses "spread()" to combine observations that were previously spread over two rows into a single row.
```{r}
table2 %>%
    spread(key = type, value = count)
```

#### 12.3 Questions

**2:** Why does this code fail? Fix it so it works.


The code fails because 1999 and 2000 are non-syntatic names and should be in backticks. The code below has been corrected so that the names are in backticks.
```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
#> Error in inds_combine(.vars, ind_list): Position must be between 0 and n
```

That is all for Chapter 12. On to the last chapter.


## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data Transformation](https://r4ds.had.co.nz/transform.html), sections 1-4.

Time to [get small.](https://www.youtube.com/watch?v=GOrdzCHnpw4) 

### 5.1: Introduction

Load the necessary libraries. As usual, type the examples into and run the code chunks.

```{r}
library(nycflights13)
library(tidyverse)
```


This code represents a tibble of a very large data set.
```{r}
flights
```

### 5.2: Filter rows with `filter()`

Study Figure 5.1 carefully. Once you learn the `&`, `|`, and `!` logic, you will find them to be very powerful tools.

This code returns a smaller subset of the data set by using "filter()" to select flights on January 1st.
```{r}
filter(flights, month == 1, day == 1)

```


This code saves the previous result to "jan1"
```{r}
jan1 <- filter(flights, month == 1, day == 1)
```


This code both prints and saves the results.
```{r}
(dec25 <- filter(flights, month == 12, day == 25))

```


This is code that contains an error. "=" should be replaced by "=="
```{r}
filter(flights, month = 1)
```


This is also code that contains an error. The could should include "near()"
```{r}
sqrt(2) ^ 2 == 2
1 / 49 * 49 == 1
```


This is the code that has been corrected to include "near()".
```{r}
near(sqrt(2) ^ 2,  2)
near(1 / 49 * 49, 1)
```


This code finds flights from November and December.
```{r}
filter(flights, month == 11 | month == 12)
```


This is a shorthand version of the previous code. This code selects every row where "x" = a value in "y".
```{r}
nov_dec <- filter(flights, month %in% c(11, 12))
```


This code shows two different ways to filter for flights that were not delayed by over 120 minutes.
```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```


This code represents missing values, which return "NA".
```{r}
NA > 5
10 == NA
NA + 10
NA / 2
```


This code shows that if values are unknown, we are unable to know if the values are equal.
```{r}
NA == NA
```


This code is used to determine if a value is missing.
```{r}
is.na(x)
```


In this code, the first filter excludes "FALSE" and "NA" answers. The second filter preserves "NA" answers by including "is.na(x)" in the filter.
```{r}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
filter(df, is.na(x) | x > 1)
```

#### 5.2 Questions

**1.1:** Find all flights with a delay of 2 hours or more.


This code represents flights with an arrival delay of at least 120 minutes.
```{r}
filter(flights, arr_delay >= 120)
```

**1.2:** Flew to Houston (IAH or HOU)


This code represents flights with destinations of either "IAH" or "HOU".
```{r}
filter(flights, dest %in% c("IAH", "HOU"))
```

**1.3:** Were operated by United (UA), American (AA), or Delta (DL).


This code represents flights where "carrier" = "UA", "AA", or "DL".
```{r}
filter(flights, carrier %in% c("AA", "DL", "UA"))
```

**1.4:** Departed in summer (July, August, and September).


This code represents flights that departed in month "7", "8", or "9".
```{r}
filter(flights, month %in%  c(7, 8, 9))
```

**1.5:** Arrived more than two hours late, but didn’t leave late.


This code represents flights that do not have a departure delay but have an arrival delay of at least 120 minutes.
```{r}
filter(flights, dep_delay <= 0, arr_delay > 120)
```

**1.6:** Were delayed by at least an hour, but made up over 30 minutes in flight. This is a tricky one. Do your best.


This code represents flights that have a departure delay of at least 60 minutes but made up over 30 minutes in flight, meaning the arrival delay is at least 30 minutes less than the departure delay.
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```

**1.7:** Departed between midnight and 6am (inclusive)


This code represents flights that departed between the times "2400" and "600".
```{r}
filter(flights, dep_time %% 2400 <= 600)
```

**2:** Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?


"between()" can be used to simplify code that selects for data where "x" falls between specified "y" values. This code uses "between()" and is a simplified version of the code in question 1.4. 
```{r}
filter(flights, between(month, 7, 9))
```

**3:** How many flights have a missing dep_time? What other variables are missing? What might these rows represent?


This code represents flights with no listed "dep_time". The code returns 8,255 flights. Other data is missing as well, such as "arr_time". These appear to be flights that never departed and were possibly canceled. 
```{r}
filter(flights, is.na(dep_time))
```

**4:** Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)

**Note:** For some context, see [this thread](https://blog.revolutionanalytics.com/2016/07/understanding-na-in-r.html)


NA ^ 0 = 1 because x ^ 0 = 1.
```{r}
NA ^ 0
```


"NA | TRUE" is not missing because "NA" could represent either "TRUE" or "FALSE", and in either case, this would return "TRUE". 
```{r}
NA | TRUE
```


Similar to the previous code, possible values for "NA" would return "FALSE".
```{r}
FALSE & NA
```

### 5.3 Arrange with `arrange()`


This code uses "arrange()" to change the order of the columns.
```{r}
arrange(flights, year, month, day)
```


This code uses "(desc)" to reorder a column so that it is in descending order.
```{r}
arrange(flights, desc(dep_delay))
```


This code includes missing values. Missing values are still at the end after the data is reordered.
```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, x)
arrange(df, desc(x))
```

#### 5.3 Questions

**1:** How could you use `arrange()` to sort all missing values to the start? (Hint: use is.na()). **Note:** This one should still have the earliest departure dates after the `NA`s. *Hint:* What does `desc()` do?


You can use "arrange()" to sort missing values to the start by sorting by (is.na(dep_time)).
```{r}
arrange(flights, desc(is.na(dep_time)), dep_time)
```

**2:** Sort flights to find the most delayed flights. Find the flights that left earliest. 

This question is asking for the flights that were most delayed (left latest after scheduled departure time) and least delayed (left ahead of scheduled time).


The following code finds the flights that were most delayed by sorting in descending order of departure delays, so the data will descend from the most delayed to the least delayed. The second portion of the code finds the flights that are the least delayed by using "arrange(flights, dep_delay)" to sort in ascending order of departure delays, so the data will be displayed in order from the least delayed to the most delayed. 
```{r}
arrange(flights, desc(dep_delay))
arrange(flights, dep_delay)
```

**3:** Sort flights to find the fastest flights. Interpret fastest to mean shortest time in the air.


This code finds the fastest flights by using "arrange(flights, air_time)" to arrange the flights in ascending order of air time, so the data displays flights in order of the shortest time to the longest time. 
```{r}
arrange(flights, air_time) 
```

*Optional challenge:* fastest flight could refer to fastest air speed. Speed is measured in miles per hour but time is minutes. Arrange the data by fastest air speed.


**4:** Which flights travelled the longest? Which travelled the shortest?


This code finds the flights that travelled the longest and the flights that travelled the shortest. "arrange(flights, desc(distance))" finds the flights that travelled the longest by sorting flights in descending order of distance so that the flights are displayed in order of longest distance to shortest distance. "arrange(flights, distance)" finds the flights that travelled the sortest by sorting flights in ascending order of distance so that the flights are displayed in order of shortest distance to longest distance.
```{r}
arrange(flights, desc(distance))
arrange(flights, distance)
```

### 5.4 Select columns with `select()`


This code selects for a subset of the "flights" data set by using "select()" to select for the desired variables. The first line selects for variables "year", "month", and "day". The second line selects for variables from "year" to "day". The third line selects for all variables except for those from "year" to "day".
```{r}
select(flights, year, month, day)
select(flights, year:day)
select(flights, -(year:day))
```


This code uses "rename()" to change a variable name from "tailnum" to "tail_num".
```{r}
rename(flights, tail_num = tailnum)
```


This code reorganizes the data by using "select()" to select variables to move to the beginning and using "everything()" to move the other variables so that they are displayed after the selected variables.
```{r}
select(flights, time_hour, air_time, everything())
```

#### 5.4 Questions

**1:** Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from flights. Find at least three ways.


This code contains three ways to select the variables "dep_time", "dep_delay", "arr_time", and "arr_delay". In the first line, variables are selected for individually by name. In the second line, variables between "dep_delay" and "arr_time" are selected for using "dep_delay:arr_time". The other variables are selected for individually. In the third line, variables are selected for by column number.
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, dep_time, dep_delay:arr_time, arr_delay)
select(flights, 4, 6, 7, 9)

```

**2:** What happens if you include the name of a variable multiple times in a `select()` call?


If you include the name of a variable multiple times, it is only displayed once. In this code, the variable "year" is included twice but is only displayed once.
```{r}
select(flights, year, year)
```

**3:** What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?

`vars <- c("year", "month", "day", "dep_delay", "arr_delay")`


"one_of()" can be used to select for variables stored as a vector. In this code, "select(flights, one_of(vars))" is used to select for variables "year", "month", "day", "dep_delay", and "arr_delay", which are stored as "vars".
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```

**4:** Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

`select(flights, contains("TIME"))`


The result of running the code may be surprising to some because the default behavior is to ignore case. "ignore.case = FALSE" can be used to change the default.
```{r}
select(flights, contains("TIME"))
select(flights, contains("TIME", ignore.case = FALSE))
```
