---
title: "HW04 Part 1: Complete Chapter 3"
author: Madison Hongsermeier-Baxter
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3

- Change "your name" in the YAML header above to your name.

### Section 3.1

Load the tidyverse packages. I've entered the first code chunk for you. **You must enter all subsequent code chunks and run the code.** 

```{r}
library("tidyverse")
```

If you get an error, then you did not install the `tidyverse` package like you were supposed to in the previous assignment. Go back to that assignment and complete that step. (This should not be the case, though.)

### Section 3.2: First steps

Enter your code chunks for Section 3.2 here.

Get in the habit of entering a brief description of what each chunk of code does. Enter this above the chunk, like I did in Section 3.1.



This code plots the mpg data frame. It plots mpg with the variable "displ" on the x axis and the variable "hwy" on the y axis.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```


This is a code template. The bracketed sections can be replaced with the relevant data in order to make graphs using ggplot2.

```{r}
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

```




#### Section 3.2 Questions

Answer the questions *completely.* Some answers will require you to write code in a code chunk and also to type text before or after the chunk. Other answers will require only code or only text. 

**1:** Run ggplot(data = mpg). What do you see?


the code "ggplot(data = mpg)" creates an empty graph.

```{r}
ggplot(data = mpg)
```



**2:** How many rows are in `mpg`? How many columns? **Hint:** Use the `dim()` function. Type `dim()` in the console to learn what `dim()` does.


"mpg" has 234 rows and 11 columns.

```{r}
dim(mpg)

```



**3:** What does the `drv` variable describe? Read the help for `?mpg` to find out.



The "drv" variable in "mpg" describes if a car is front-wheel drive, rear wheel drive, or four-wheel drive.



**4:** Make a scatterplot of `hwy` vs `cyl`.



This is a scatterplot of hwy vs cyl.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))
```



**5:** What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?


This a scatterplot of class vs drv. The plot is not useful because the correlation between the variables "class" and "drv" is not relevant to the relationship between engine size and fuel efficiency. Also, the plot is not useful because it does not show any clear correlation between the variables.


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))
```



### Section 3.3: Aesthetic mappings



This code creates a scatterplot of "hwy" vs "displ" with different colored points to represent different classes of vehicle.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```




This code creates a scatterplot of "hwy" vs "displ" with different sized points to represent different classes of vehicle.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```





This code creates a scatterplot of "hwy" vs "displ" with different transparencies of points to represent different classes of vehicle.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```




This code creates a scatterplot of "hwy" vs "displ" with different shapes of points to represent different classes of vehicle.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```



This code creates a scatterplot of "hwy" vs "displ" where all of the points are blue.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```



#### Section 3.3 questions

**1:** What’s gone wrong with this code? Why are the points not blue?


The original code does not result in blue points because "color = blue" was within "aes()" The code block below has a corrected version of the code and results in a graph with blue points.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```



**2:** Which variables in `mpg` are categorical? Which variables are continuous? (*Hint:* type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg?


The variables "model", "trans", "drv", "fl", and "class" are categorical. The variables "hwy", "cty", "cyl", "year", and "displ" are continuous. When you run mpg, you can see which varibles are categorical and which variables are continous because the  continous variables of mpg are numerical and have a logical order, and the categorical variables have a limited number of categories and no logical order.



**3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **Note:** You need several code chunks here to answer this question.



When mapping a categorical variable, such as "class", to a color, it results in a finite number of different unordered colors to represent the values. However, when mapping a continuous variable, such as "cty", to a color, the values are represented by a color spectrum. The spectrum has a logical order and infinite different colors.


color = class
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

color = cty
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cty))
```






When mapping a categorical variable, such as "class", to a size, it results in a finite number of different unordered sizes to represent the values. When mapping a continuous variable, such as "cty", to a size, the size of the points increase and decrease according to the logical order of the values.


size = class
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```


size = cty
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = cty))
```






When mapping a categorical variable, such as "class", to a shape, it results in a finite number of different unordered shapes to represent the values. However, when mapping a continuous variable, such as "cty", to a shape, it results in the error message "A continuous variable can not be mapped to shape". The shapes are not able to properly convey the nature of the continous variable because the shape aesthetic is finite and does not have a logical order.


shape = class
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```


shape = cty
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = cty))
```


**4:** What happens if you map the same variable to multiple aesthetics?



If you map the same variable to multiple aesthetics, the aesthetics are applied to the variable simultaneously. For example, in the graph below, each class is associated both with a color and with a shape.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class, color = class))
```


**5:** What does the stroke aesthetic do? What shapes does it work with? (*Hint:* use `?geom_point`)


The stroke aesthetic can be applied to shapes that have a border and is used to change the width of the border.



**6:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?


When mapping the aesthetic color to "displ < 5" it results in a graph where all of the values < 5 are assigned to one color and all of the values > 5 are assigned to a second color, visually dividing the data into the two different categories.


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))
```



### Section 3.5: Facets


This code splits up mpg into facets that display subsets of the data.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```




This code facets the plot on the combinaion of the vaiables "drv" and "cyl".

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```






#### Section 3.5 questions

**1:** What happens if you facet on a continuous variable? **Note:** Write a prediction, then make and run your code chunk.



Prediction: The facets will be organized according to a logical order.


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ cty, nrow = 2)
```





**2:** What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?


There are empty cells in the plot because there are values at which the two variables do not coincide. For example, none of the vehicles with 4 cylinders have rear-wheel drive, so there is an empty cell at the intersection of these 2 variables. The following code shows where values of the 2 variables coincide and can be used to explain where the empty cells are.


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```




**3:** What plots does the following code make? What does `.` do?

The code results in one plot faceted by the variable "drv" and another plot faceted by the variable "cyl". The "." is used to determine whether the plot is faceted in the rows dimension or in the columns dimension. For example, in the first plot, there is a "." in the column section of the code, specifying that the data should be faceted in the rows dimension.


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```






**4:** Take the first faceted plot below. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?


An advantage of faceting over color aesthetic is that you can more easily compare the "hwy" vs "displ" scatterplots for each class. An advantage of using the color aestheic over faceting is that you can more easily compare the values of "hwy" and "displ".


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```





**5:** Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have nrow and ncol argument?


"nrow" is used to specify the number of rows. "ncol" is used to specify the number of columns. There are other options that control the layout of individual panels, such as "scales" and "dir". "facet_grid()" doesn't have nrow and ncol arguments because the numbers of rows and columns are already determined by the facets and directions.



**6:** When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?


The variable with the more unique levels should be in the columns so that the plot will have a more favorable layout and be easier to read. 



### Section 3.6: Geometric objects



#### Section 3.6.1 Questions

**1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart? **Note:** You will have to make some types of these charts in Part 2 of this assignment.


**2:** Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.



**3:** What does `show.legend = FALSE` do? What happens if you remove it?
**Note:** Skip the "Why do you think I used it earlier in the chapter?" question.


**4:** What does the `se` argument to `geom_smooth()` do?


**5:** Will these two graphs look different? Why/why not?


**6:** Recreate the R code necessary to generate the following graphs. **Note:** Enter a separate code chunk for each graph, numbered 1 to 6, starting row-wise at upper left. 

1: upper left.


2: upper right.


3: middle left.


4: middle right.


5: lower left.


6: lower right. *Hint:* Read the help file for `geom_point`. Use `shape = 21`, `stroke = 2`, and `size = 3`.

### Section 3.7: Statistical transformations.

**Skip this section.** We may return to it later.


### Section 3.8 Position adjustments


#### Section 3.8 Questions

**1:** What is the problem with this plot? How could you improve it. *Hint:* Remember the size of the `mpg` data set from earlier in this exercise. Does the number of points seem correct? **Note:** Write the code to improve it. 


**2:** What parameters to `geom_jitter()` control the amount of jittering?


**3:** Compare and contracts `geom_jitter()` with `geom_count()`.


**4:** What is the default position adjustement for `geom_boxplot()`? Create a visualization of the `mpg` dataset that demonstrates it.


### Section 3.9: Coordinate systems

**Note:** Read about and do only the code chunk associated with `coord_flip()`. `coord_flip()` is a great layer to use when you have categorical data with longish names as it makes the text easier to read.

### Section 3.10: The layered grammar of graphics

Just read this section as a contextual review of everything you just learned.


