---
title: "HW04 Part 2: Graph some biological data"
author: Madison Hongsermeier-Baxter
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r}
library("tidyverse")
```

Plot 1

    Plot type: scatterplot
    Dataset: trees
    x-axis: height
    y-axis: girth

#### Plot 1

```{r}
data(trees)
```

```{r}
?trees
```


This is a scatterplot of tree height vs girth. The scatterplot shows a positive correlation between height and girth. In general, as height increases, girth increases as well.
```{r}
ggplot(data = trees) + 
  geom_point(mapping = aes(x = Height, y = Girth))
```

Plot 2

Apply some of your skills that you learned during Assignment 02. You will make two vectors, then combine them into a data frame for plotting. Review the assignments if necessary.

    Make a vector called year for 1821 to 1934. Rememeber how to use : to make a sequence of numbers?

    Look at the class() of the lynx dataset. The lynx dataset is a “time series” class (ts). You can convert the time series data to a vector by using the as.vector() function. Just put the dataset name inside the parentheses. Assign this to the variable pelts.

    Make a dataframe called lynx_pelts from these two vectors.
    Plot type: linegraph
    Dataset: lynx_pelts (see above)
    x-axis: year
    y-axis: pelts

    Make the line color maroon. Maroon is one of the default R colors.

#### Plot 2

```{r}
data(lynx)
```

```{r}
?lynx
```

```{r}
year <- 1821:1934
```

```{r}
class(lynx)
```

```{r}
pelts <- as.vector(lynx)
```

```{r}
lynx_pelts <- data.frame(year, pelts)
```

```{r}
lynx_pelts
```


This is a line graph of "year" vs "pelts". The graph shows a pattern where "pelts" fluctuates over time. There appears to be a repeating cycle of a larger fluctuation followed by three smaller fluctuations. The cycle takes place over approximately 40 years then repeats. 
```{r}
ggplot(data = lynx_pelts) + 
  geom_line(mapping = aes(x = year, y = pelts), color = "maroon")
```

Plot 3

    Plot type: scatterplot
    Dataset: iris
    x-axis: petal length
    y-axis: petal width
    Point color and shape by species. You do not have to use fillable shapes.
    Point size of 2
    Add a labs layer to change the x- and y-axis labels so that they do not have periods in the names (i.e., Petal Length, Petal Width).

#### Plot 3

```{r}
data(iris)
```

```{r}
?iris
```


This is a scatterplot of petal length vs petal width. The scatterplot shows a positive correlation between length and width. In general, as length increases, width increases as well.
```{r}
ggplot(data = iris) + 
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, color = Species, shape = Species)) +
  labs(x = "Petal Length", y = "Petal Width")
```

Plots 4 and 5

    This requires two code chunks, which will be nearly identical
    Plot type: Violin plot (look up geom_violin)
    Dataset: msleep
    x-axis: vore
    y-axis: sleep_rem
    fill = grayXX where XX is either 30 or 70.

    In your description, describe in your own words what violin plots display (you can search the interwebs), and what is the difference among the two versions of gray shading. Hint: the grays extend from gray0 to gray100. You can learn more about colors in R from this PDF file.

#### Plot 4 and 5

```{r}
data(msleep)
```

```{r}
?msleep
```


This is a violin plot of "vore" vs "sleep_rm". A violin plot is similar to a box plot because both are used to represent distribution of data. However, violin plots use shapes to more accurately portray the variations in the data. The first violin plot is filled with the color gray30, and the second violin plot is filled with the color gray70. Grays extend from gray0 to gray100, with gray100 being the lightest and gray 0 being the darkest, so the gray70 plot is lighter than the gray30 plot.
```{r}
ggplot(data = msleep) + 
  geom_violin(mapping = aes(x = vore, y = sleep_rem), fill = "gray30")
```

```{r}
ggplot(data = msleep) + 
  geom_violin(mapping = aes(x = vore, y = sleep_rem), fill = "gray70")
```

Plot 7

    Plot type: boxplot
    Dataset: msleep
    x-axis: order
    y-axis: sleep_total
    use coord_flip()

#### Plot 7


These are boxplots of "order" vs "sleep_total". In the second plot, "coord_flip" has been used to switch the x axis and y axis. This organizes the data in a way that makes it easier to visualise the data and read the "order" axis.
```{r}
ggplot(data = msleep) + 
  geom_boxplot(mapping = aes(x = order, y = sleep_total))

ggplot(data = msleep) + 
  geom_boxplot(mapping = aes(x = order, y = sleep_total)) +
  coord_flip()
```


Plot 8

    Plot type: boxplot with points
    Dataset: msleep
    x-axis: conservation
    y-axis: awake
    You must have a boxplot layer, a point layer, and a jitter layer.
    color by conservation status.
    You may use coord_flip but it is not required. Try both and choose the one you think looks best.
    Add a lab layer to change both axis labels so each starts with an upper-case letter.
    Search the internet to find out how to change the legend title to Conservation. Make that change. (Do not try to change the actual legend entries like “cd” and “vu”). Note: This can be done a couple of different ways but scale_color_discrete() is one good way.

#### Plot 8


This plot for "conservation" vs "awake" utilizes both a boxplot layer and a scatterplot layer to display the data. The different values for "conservation" are also specified by color. 
```{r}
ggplot(data = msleep, mapping = aes(x = conservation, y = awake, color = conservation)) + 
  geom_boxplot() + 
  geom_point() +
  geom_jitter() +
  labs(x = "Conservation", y = "Awake") +
  labs(color= "Conservation")  
```

Plots 9 and 10.

    Make two scatterplots of your choice, with the following constraints.
    One should plot any one of the “sleep” or “awake” variables against body weight. The other should plot any one of the “sleep” or “awake” variables against brain weight.
    In each plot, color the points by one of the nominal data categories. Use a different category for each plot.
    Apply facet_wrap() to at least one of the plots using one of the nominal variables. You decide whether you use 2 or 3 columns. Hint: use one of the nominal variables with relatively few different types for wrapping. Explore: What happens if you use a nominal variable like genus, with lots of types?
    You should try a few versions of each graph until you find combinations that you think show some clear trends.

#### Plot 9 and 10


This is a scatterplot of "awake" vs "bodywt". The plot shows that the animals with the highest weights are also awake for the longest durations. "color = vore" shows that most of these largest animals are also herbivores.
```{r}
ggplot(data = msleep, mapping = aes(x = awake, y = bodywt, color = vore)) + 
  geom_point()
```


This is a scatterplot of "sleep_total" vs "brainwt". The plot shows that animals with higher brain weights appear to get less total sleep. Also, "facet_wrap(vore)" shows that "herbi" have some of the highest brain weights and "insect" and "NA" have the lowest overall brain weights.
```{r}
ggplot(data = msleep, mapping = aes(x = sleep_total, y = brainwt, color = order)) + 
  geom_point() +
  facet_wrap(~ vore, nrow = 2)
```
